---
title: "absenteesim_at_work"
author: "aria_fredman"
date: "10/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

loading packages
```{r loading libraries}

library(CausalImpact); library(tidyverse)

```

setting the working directory
reading in the csv data
```{r wd reading data, echo=FALSE}

setwd("~/Documents/data_science/data/absent")
df <- read.csv("absent.csv")

```

looking at the first few rows
```{r 1st head}

head(df)

```
That looks very wrong
probably a parsing issue
```{r sep}

df <- read.csv("absent.csv", sep = ";")
head(df)

```
much better!
better look at column names
```{r }

colnames(df)

```
I want underscores instead of periods
all underscore
```{r }

colnames(df) <-
  tolower(
#taking all periods in column names and subbing with underscores
#since . is a special character meaning any character, we need to escape it first
    gsub("\\.", "_", colnames(df)) 
  )

colnames(df)

```
Feature descriptions:
1. Individual identification (ID)
2. Reason for absence (ICD).
Absences attested by the International Code of Diseases (ICD) stratified into 21 categories (I to XXI) as follows:

I Certain infectious and parasitic diseases
II Neoplasms
III Diseases of the blood and blood-forming organs and certain disorders involving the immune mechanism
IV Endocrine, nutritional and metabolic diseases
V Mental and behavioural disorders
VI Diseases of the nervous system
VII Diseases of the eye and adnexa
VIII Diseases of the ear and mastoid process
IX Diseases of the circulatory system
X Diseases of the respiratory system
XI Diseases of the digestive system
XII Diseases of the skin and subcutaneous tissue
XIII Diseases of the musculoskeletal system and connective tissue
XIV Diseases of the genitourinary system
XV Pregnancy, childbirth and the puerperium
XVI Certain conditions originating in the perinatal period
XVII Congenital malformations, deformations and chromosomal abnormalities
XVIII Symptoms, signs and abnormal clinical and laboratory findings, not elsewhere classified
XIX Injury, poisoning and certain other consequences of external causes
XX External causes of morbidity and mortality
XXI Factors influencing health status and contact with health services.

And 7 categories without (CID) patient follow-up (22), medical consultation (23), blood donation (24), laboratory examination (25), unjustified absence (26), physiotherapy (27), dental consultation (28).
3. Month of absence
4. Day of the week (Monday (2), Tuesday (3), Wednesday (4), Thursday (5), Friday (6))
5. Seasons (summer (1), autumn (2), winter (3), spring (4))
6. Transportation expense
7. Distance from Residence to Work (kilometers)
8. Service time
9. Age
10. Work load Average/day
11. Hit target
12. Disciplinary failure (yes=1; no=0)
13. Education (high school (1), graduate (2), postgraduate (3), master and doctor (4))
14. Son (number of children)
15. Social drinker (yes=1; no=0)
16. Social smoker (yes=1; no=0)
17. Pet (number of pet)
18. Weight
19. Height
20. Body mass index
21. Absenteeism time in hours (target) 

```{r }

View(df %>% arrange(id, month_of_absence, day_of_the_week))

```

there doesn't seem to be a way to arrange the data in order

```{r }



```

```{r }



```

```{r }



```

```{r }



```

```{r }



```

