---
title: "causal_inferencing"
author: "aria_fredman"
date: "6/18/2020"
output: html_document
---
playing with causal inferencing

loading libraries of interest
```{r}
library(tidyverse); library(MatchIt); library(mosaic); library(psych)
```

setting the working directory
```{r}
setwd("C:/Users/Aria/Documents/ds_practice/R/causal_inferencing")
```


loading the lalonde data from the MatchIt library
```{r}
data(lalonde)
```
A data frame with 614 observations (185 treated, 429 control). There are 10 
variables measured for each individual. "treat" is the treatment assignment 
(1=treated, 0=control). "age" is age in years. "educ" is education in number of 
years of schooling. "black" is an indicator for African-American 
(1=African-American, 0=not). "hispan" is an indicator for being of Hispanic 
origin (1=Hispanic, 0=not). "married" is an indicator for married (1=married, 
0=not married). "nodegree" is an indicator for whether the individual has a high 
school degree (1=no degree, 0=degree). "re74" is income in 1974, in U.S. dollars. 
"re75" is income in 1975, in U.S. dollars. "re78" is income in 1978, in U.S. dollars.


eyeballing the first few rows
```{r}
df_l <- lalonde

head(lalonde)
```

examing the data structure
```{r}
str(df_l)
```
column names also all look fine

curious mostly whether we have exactly 2 levels in the variables where we'd expect
that
```{r}
df_l %>% 
  summarise_all(n_distinct)
```

statistics for the different columns
```{r}
psych::describe(df_l)
```

statistics for the different columns by treatment
```{r}
psych::describeBy((df_l %>%
                   select(-age)), 
                df_l$treat)

```
we see the classes are pretty imbalanced with far fewer in the treatment condition
more blacks, less latinos, less married and more people without degrees in the
treatment condition
salaries are lower for each of those years, but seeing that the confounding variables
don't seem to be equally spread across the conditions it's unclear that the 
treatment didn't work
this could be especially the case if we're interested in whether the treatment
increases earnings relative to a starting point, and not over all
the people in the treatment condition had a lower starting point so even if 
their end point is lower they may still have come out ahead
either way we want to compare apples to apples across conditions and here we're
comparing apples to pears across conditions

i want to create two outcome variables
a 
```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```
