---
title: "daily_demand_forecasting_orders"
author: "aria_fredman"
date: "10/13/2020"
output: html_document
---
setting working directory
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "~/Documents/data_science/data/daily_demand_orders")

```

loading packages
```{r loading libraries}

library(CausalImpact); library(tidyverse); library(magrittr)

```

setting the working directory
reading in the csv data
NOTE all data were collected during a single month
```{r wd reading data, echo=FALSE}

setwd("~/Documents/data_science/data/daily_demand_orders")
df <- read.csv("daily_demand_forecasting_orders.csv")

```

looking at the first few rows
```{r 1st head}

head(df)

```
That looks very wrong
probably a parsing issue
```{r sep}

setwd("~/Documents/data_science/data/daily_demand_orders")
df <- read.csv("daily_demand_forecasting_orders.csv", sep = ";")
head(df)

```
much better!
better look at column names
```{r }

colnames(df)

```
I want to remove periods at the end of the name
underscores instead of periods
all underscore
```{r }

colnames(df) <-
  gsub("[[:punct:]]+$", "", colnames(df))

#looking for all punctuation with nothing after it and replacing it with nothing

colnames(df)

```

I want to replace one or more periods
with underscores instead of periods
all underscore letters
```{r }

colnames(df) <-
  tolower(
#taking all one or more periods in column names and subbing with underscores
#since . is a special character meaning any character, we need to escape it first
    gsub("\\.+", "_", colnames(df)) 
  )

colnames(df)

```
some of these names are absurd and need to be shortened
I don't want columns 8 through 12
```{r }

colnames(df)[1] <-
  "week"

colnames(df)[2] <-
  "day"

df %<>% select(-8:-12)

colnames(df)

```


```{r }



```

```{r }



```

```{r }



```

```{r }



```

```{r }



```

